---
title: "HoshiMap"
author: "Lucas Wautot"
format: html
editor: visual
---

# HoshiMap : Finding the best stargazing spots throughout Japan

***GitHub** :* *https://github.com/LucasWtt/Hoshimap*

***Google Drive Folder** : https://drive.google.com/drive/folders/1iwF9XNg6ltKaJH5gH9Hwz0whuSO3HgM7*

The objective of this report is to identify and highlight the promising places in any Japanese prefecture to perform astronomical observations.

The report is divided in two main parts. The first one describe the steps taken from data aquisition to cloud storage creation.

The second half of this work encompasses the code linked to the mapping tool itself, which can be used by anyone to identify high-potential stargazing spots.

## RStudio Code (part 1) : From raw data to cloud-available refined data

To measure the potential of the Japanese territory, four main factors where taken into account :

**Light pollution** : Human activities have enormous impact on night time brightness. Cities and industrial operations transformed nocturnal landscape and strongly decreased the average quality of stargazing activities. Urban-oriented environments are highly affected by light pollution.

**Land use** : Different territory occupations present opportunities or limitations in regards to astronomical observations. Terrain environments such as forests or urban cities allow for less possibilities compared to croplands or bare soil.

**Sky openness** : In a country with a strong relief is less convenient than flat terrain to observe the horizon. Due to its active geomorphological processes, Japanese landscape is strongly impacted by this. Some valleys might have high quality sky darkness, but have a high proportion of its sky being covered by mountainside, restricting the window of observation.

**Accessibility** : a

**Aerosol optical thickness** : a

**Hillshade** :

**Altitude and Slope** : Despite being less impactful than the other variables, altitude also influences the quality of stargazing. Atmospheric perturbations have a strong influence on telescope related activities. At an altitude of 1500 meters, the atmospheric pressure is only 85% of what it is at sea level. Moreover, lower atmosphere is generally turbulent due to the proximity with the ground and its vegetation. These factors can theoretically influence visibility up to 20%. This variable acts as a proxy for atmospheric stability.

Slope, even though it is highly related to altitude, has a completely different role in this field. It doesn't have a direct relationship with the atmosphere, but with the ground. Steep slope environments are generally viewed as noxious for astronomy related activities, instruments such as telescopes require a relatively flat area to be used.

Despite their relationships with stargazing, altitude and slope have not been utilized in the version of HoshiMap released here. This is because these two variables are less vital to the subject, and limiting the number of variables taken into the computational power drastically increases the speed of rendering of HoshiMap. Slope itself appeared to be strongly correlated with sky openness. However, the data for both of those parameters is available in the Google Drive Folder.

## Preprocessing code

```{r}
# ==============================================================================
# PACKAGES & SETUP
# ==============================================================================
if (!require("pacman")) install.packages("pacman")
pacman::p_load(terra, sf, dplyr, stringr, googledrive, ggplot2, tidyterra)

# Create output structure
base_dirs <- c("DEM", "Landuse", "Lightpollution", "Accessibility", 
               "OpticalDepth", "Hillshade", "Slope", "Skyview")
lapply(paste0("output/", base_dirs, "_prefectures"), dir.create, recursive = TRUE, showWarnings = FALSE)

# ==============================================================================
# DATA IMPORT & INITIAL CLEANING
# ==============================================================================
# Load Country Boundary
JP_Shp <- vect("data/JP_Shp/jp.shp")

# Load and Clean Rasters
JP_OpticalDepth  <- rast("data/Aerosol_Optical_Thickness.tif")
JP_OpticalDepth[JP_OpticalDepth == -9999] <- NA
JP_OpticalDepth  <- JP_OpticalDepth / 10000

JP_Accessibility <- rast("data/JP_cost/accessibility_to_cities_2015_v1.0.tif")
JP_Accessibility[JP_Accessibility == -9999] <- NA

# Combine Landuse Tiles
lu_list <- list.files("data/landuse_dataset", pattern = "\\.tif$", full.names = TRUE)
JP_Landuse <- do.call(merge, lapply(lu_list, rast))

# Load others
JP_DEM            <- rast("data/JP_DEM.tif")
JP_Skyview        <- rast("data/JP_Skyview.tif")
JP_Lightpollution <- rast("data/JP_Lightpollution")

# ==============================================================================
# CROP & MASK TO NATIONAL BOUNDARY
# ==============================================================================
# This standardizes the "Working Area" before splitting into prefectures
layers <- list(
  DEM             = JP_DEM,
  Skyview         = JP_Skyview,
  Landuse         = JP_Landuse,
  Lightpollution  = JP_Lightpollution,
  Accessibility   = JP_Accessibility,
  OpticalDepth    = JP_OpticalDepth
)

# Apply national crop/mask to all layers
layers <- lapply(layers, function(r) {
  r_crop <- crop(r, JP_Shp)
  mask(r_crop, JP_Shp)
})

# ==============================================================================
# PREFECTURE SPLITTING LOOP
# ==============================================================================
for(i in 1:nrow(JP_Shp)) {
  
  pref      <- JP_Shp[i,]
  pref_name <- gsub("[^[:alnum:]_]", "_", pref$name[1])
  
  if(pref_name == "Okinawa") next # Skipping due to coordinate mismatch
  
  message(paste("--- Processing Prefecture:", pref_name, "---"))
  
  # 1. Create Individual Prefecture Layers
  for(lyr_name in names(layers)) {
    
    # Crop and Mask to Prefecture
    p_crop <- crop(layers[[lyr_name]], pref)
    p_mask <- mask(p_crop, pref)
    
    # Save standard layers
    out_path <- paste0("output/", lyr_name, "_prefectures/", lyr_name, "_", pref_name, ".tif")
    writeRaster(p_mask, out_path, overwrite = TRUE, gdal = c("COMPRESS=LZW"))
    
    # 2. Generate Derived Terrain Layers (only if current layer is DEM)
    if(lyr_name == "DEM") {
      # Slope
      slp <- terrain(p_mask, v = "slope", unit = "degrees")
      writeRaster(slp, paste0("output/Slope_prefectures/Slope_", pref_name, ".tif"), overwrite = TRUE)
      
      # Hillshade
      asp <- terrain(p_mask, v = "aspect", unit = "radians")
      slp_rad <- terrain(p_mask, v = "slope", unit = "radians")
      hld <- shade(slp_rad, asp, angle = 45, direction = 315)
      writeRaster(hld, paste0("output/Hillshade_prefectures/Hillshade_", pref_name, ".tif"), overwrite = TRUE)
    }
  }
}

# ==============================================================================
# ALIGNMENT (RESAMPLING)
# ==============================================================================
# Ensure all rasters have the exact same grid as the DEM for Shiny math
for (i in 1:nrow(JP_Shp)) {
  pref_name <- gsub("[^[:alnum:]_]", "_", JP_Shp$name[i])
  if(pref_name == "Okinawa") next
  
  # Reference Raster (DEM)
  ref <- rast(paste0("output/DEM_prefectures/DEM_", pref_name, ".tif"))
  
  # Resample Continuous Data (Bilinear)
  to_resample_cont <- c("Lightpollution", "Accessibility", "OpticalDepth")
  for(lyr in to_resample_cont) {
    path <- paste0("output/", lyr, "_prefectures/", lyr, "_", pref_name, ".tif")
    if(file.exists(path)) {
      r <- rast(path)
      r_aligned <- resample(r, ref, method = "bilinear")
      writeRaster(r_aligned, path, overwrite = TRUE, gdal = c("COMPRESS=LZW"))
    }
  }
  
  # Resample Categorical Data (Nearest Neighbor)
  lu_path <- paste0("output/Landuse_prefectures/Landuse_", pref_name, ".tif")
  if(file.exists(lu_path)) {
    r_lu <- rast(lu_path)
    lu_aligned <- resample(r_lu, ref, method = "near")
    writeRaster(lu_aligned, lu_path, overwrite = TRUE, gdal = c("COMPRESS=LZW"))
  }
}

# ==============================================================================
# EXPORT INDIVIDUAL SHAPEFILES
# ==============================================================================
shp_sf <- st_as_sf(JP_Shp)
dir.create("output/Shapefile_prefectures", showWarnings = FALSE)

for (i in 1:nrow(shp_sf)) {
  name_clean <- gsub("[^[:alnum:]_]", "_", shp_sf$name[i])
  st_write(shp_sf[i, ], 
           paste0("output/Shapefile_prefectures/Shp_", name_clean, ".shp"), 
           delete_dsn = TRUE, quiet = TRUE)
}

message("All preprocessing completed successfully.")
```

## RStudio Code (part 2) : Open source tool

Now that the data have been cleaned, the following code is composed of two chunks

Despite their relationships with stargazing, altitude and slope have not been utilized in the version of HoshiMap released here. This is because these two variables are less vital to the subject, and limiting the number of variables taken into the computational power drastically increases the speed of rendering of HoshiMap. Slope itself appeared to be strongly correlated with sky openness. However, the data for both of those parameters is available in the Google Drive Folder.

## HoshiMap - Cloud Version

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(terra, sf, leaflet, shiny, shinythemes, bslib, googledrive)

# ====================== PUBLIC DRIVE SETUP =====================
drive_deauth() 
drive_root_id <- as_id("1iwF9XNg6ltKaJH5gH9Hwz0whuSO3HgM7")

# ===================== PERFORMANCE SETTINGS ====================
VIS_FACT <- 4

# ========================== FUNCTIONS ==========================

get_public_drive_file <- function(subfolder_name, file_name) {
  root_content <- drive_ls(drive_root_id)
  target_folder <- root_content[root_content$name == subfolder_name, ]
  if(nrow(target_folder) == 0) return(NULL)
  
  sub_content <- drive_ls(as_id(target_folder$id[1]))
  target_file <- sub_content[sub_content$name == file_name, ]
  if(nrow(target_file) == 0) return(NULL)
  
  temp_path <- file.path(tempdir(), file_name)
  drive_download(as_id(target_file$id[1]), path = temp_path, overwrite = TRUE)
  return(temp_path)
}

load_raster_drive <- function(folder, prefix, pref) {
  fname <- paste0(prefix, pref, ".tif")
  path <- get_public_drive_file(folder, fname)
  if (is.null(path)) stop(paste("File not found:", fname))
  rast(path)
}

load_hillshade_drive <- function(folder, prefix, pref) {
  fname <- paste0(prefix, pref, ".tif")
  path <- get_public_drive_file(folder, fname)
  if (is.null(path)) return(NULL)
  rast(path)
}

load_contour_drive <- function(folder, prefix, pref) {
  extensions <- c(".shp", ".shx", ".dbf", ".prj")
  base_name <- paste0(prefix, pref)
  local_shp_path <- ""
  
  for (ext in extensions) {
    fname <- paste0(base_name, ext)
    p <- get_public_drive_file(folder, fname)
    if (ext == ".shp") local_shp_path <- p
  }
  
  if (local_shp_path == "" || !file.exists(local_shp_path)) return(NULL)
  st_read(local_shp_path, quiet = TRUE)
}

landuse_mask <- function(r) {
  exclude <- c(1,2,6,7,8,9,11)
  !(r %in% exclude)
}

# ============================ UI ===============================

ui <- fluidPage(
  theme = bs_theme(
    bg = "#0B0C10",
    fg = "#FFFFFF",
    primary = "#FFD700",
    base_font = font_google("Orbitron")
  ),
  titlePanel("🌌 HoshiMap | Cloud Version"),
  sidebarLayout(
    sidebarPanel(
      selectInput("pref", "Prefecture", choices = NULL),
      sliderInput("access", "Max Travel Time (min)", 0, 120, 45),
      sliderInput("optical", "Max Optical Depth", 0, 5, 0.5),
      sliderInput("light", "Max Light Pollution", 0, 100, 30),
      sliderInput("sky", "Min Sky View", 0, 1, 0.5)
    ),
    mainPanel(leafletOutput("map", height = 700))
  )
)

# ============================ SERVER ===========================

server <- function(input, output, session) {
  
  # Populate Prefectures
  observe({
    acc_folder <- drive_ls(drive_root_id, pattern = "Accessibility_prefectures")
    if(nrow(acc_folder) > 0) {
      files <- drive_ls(as_id(acc_folder$id[1]), pattern = "\\.tif$")
      prefs <- gsub("Accessibility_|\\.tif", "", files$name)
      updateSelectInput(session, "pref", choices = prefs)
    }
  })
  
  # Observatories Data
  observatories_data <- reactive({
    root_files <- drive_ls(drive_root_id)
    csv_item <- root_files[root_files$name == "Observatories.csv", ]
    if(nrow(csv_item) > 0) {
      temp_csv <- file.path(tempdir(), "Observatories.csv")
      drive_download(as_id(csv_item$id[1]), path = temp_csv, overwrite = TRUE)
      df <- read.csv(temp_csv, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
      df$Type <- trimws(df$Type)
      df <- df[!is.na(df$Longitude) & df$Longitude != "", ]
      df$Longitude <- as.numeric(df$Longitude)
      df$Latitude  <- as.numeric(df$Latitude)
      return(df)
    }
    return(data.frame())
  })
  
  raster_cache <- reactiveVal(list())
  contour_cache <- reactiveVal(list())
  
  get_pref_data <- function(pref) {
    cache <- raster_cache()
    if (!pref %in% names(cache)) {
      showNotification(paste("Streaming terrain and data for", pref), type = "message")
      cache[[pref]] <- list(
        landuse   = load_raster_drive("Landuse_prefectures", "Landuse_", pref),
        access    = load_raster_drive("Accessibility_prefectures", "Accessibility_", pref),
        opt       = load_raster_drive("OpticalDepth_prefectures", "OpticalDepth_", pref),
        light     = load_raster_drive("Lightpollution_prefectures", "Lightpollution_", pref),
        sky       = load_raster_drive("Skyview_prefectures", "Skyview_", pref),
        hillshade = load_hillshade_drive("Hillshade_prefectures", "Hillshade_", pref)
      )
      raster_cache(cache)
    }
    cache[[pref]]
  }

  get_pref_contour <- function(pref) {
    cache <- contour_cache()
    if (!pref %in% names(cache)) {
      cache[[pref]] <- load_contour_drive("Shapefile_prefectures", "Shp_", pref)
      contour_cache(cache)
    }
    cache[[pref]]
  }

  # ========================= ANALYSIS ==========================
  result_rasters <- reactive({
    req(input$pref)
    r <- get_pref_data(input$pref)
    
    mask_lu <- landuse_mask(r$landuse)
    valid <- mask_lu & (r$access <= input$access) & (r$opt <= input$optical) & 
             (r$light <= input$light) & (r$sky >= input$sky)
    
    res <- mask(r$access, valid, maskvalues = 0)
    
    # Process Hillshade
    hill <- r$hillshade
    if (!is.null(hill)) {
      hill <- aggregate(hill, fact = VIS_FACT, fun = mean, na.rm = TRUE)
    }
    
    list(
      result = aggregate(res, fact = VIS_FACT, fun = mean, na.rm = TRUE),
      hillshade = hill
    )
  })

  # ============================ MAP ============================
  output$map <- renderLeaflet({
    req(input$pref)
    
    data_list <- result_rasters()
    res_viz <- data_list$result
    hill_viz <- data_list$hillshade
    
    contour <- get_pref_contour(input$pref)
    obs <- observatories_data()
    obs_sub <- subset(obs, Prefecture == input$pref)
    
    obs_types <- c("Optical / Infrared", "Radio / Astronomical")
    obs_colors <- c("#5B21B6", "#38BDF8")
    obs_pal <- colorFactor(palette = obs_colors, domain = obs_types, na.color = "#808080")
    
    pal <- colorNumeric("YlGnBu", domain = values(res_viz), na.color = "transparent")

    m <- leaflet() %>%
      addProviderTiles("CartoDB.DarkMatter")
    
    # Add Hillshade Overlay first (background)
    if (!is.null(hill_viz)) {
      m <- m %>% addRasterImage(hill_viz, opacity = 0.35, project = TRUE, colors = "Greys")
    }
    
    # Add Analysis Raster
    m <- m %>% addRasterImage(res_viz, colors = pal, opacity = 0.7, project = TRUE)
    
    # Add Boundary
    if (!is.null(contour)) {
      m <- m %>% addPolygons(
        data = contour, color = "white", weight = 2, fillOpacity = 0
      )
    }

    # Add Markers
    if (nrow(obs_sub) > 0) {
      m <- m %>% addCircleMarkers(
        data = obs_sub, lng = ~Longitude, lat = ~Latitude,
        color = ~obs_pal(Type), fillColor = ~obs_pal(Type),
        radius = 6, fillOpacity = 0.9, stroke = TRUE, weight = 1,
        popup = ~paste0("<b>", Observatory, "</b><br>", Type)
      )
    }

    m %>%
      addLegend(pal = pal, values = values(res_viz), title = "Travel Time (min)", position = "bottomleft") %>%
      addLegend(
        position = "bottomright",
        pal = obs_pal,
        values = obs_types,
        title = "Observatory Type",
        opacity = 1
      )
  })
}

shinyApp(ui, server)
```

## HoshiMap - Local Version

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(terra, sf, leaflet, shiny, shinythemes, bslib)

# ========================== BASE PATH ==========================
base_path <- "Drive"

# ===================== PERFORMANCE SETTINGS ====================
VIS_FACT <- 4

# ========================== FUNCTIONS ==========================

load_raster_local <- function(folder, prefix, pref) {
  path <- file.path(base_path, folder, paste0(prefix, pref, ".tif"))
  if (!file.exists(path)) stop(paste("Missing file:", path))
  rast(path)
}

load_hillshade_local <- function(folder, prefix, pref) {
  path <- file.path(base_path, folder, paste0(prefix, pref, ".tif"))
  if (!file.exists(path)) return(NULL)
  rast(path)
}

load_contour_local <- function(folder, prefix, pref) {
  path <- file.path(base_path, folder, paste0(prefix, pref, ".shp"))
  if (!file.exists(path)) return(NULL)
  st_read(path, quiet = TRUE)
}

landuse_mask <- function(r) {
  exclude <- c(1,2,6,7,8,9,11)
  !(r %in% exclude)
}

# ============================ UI ===============================

ui <- fluidPage(
  theme = bs_theme(
    bg = "#0B0C10",
    fg = "#FFFFFF",
    primary = "#FFD700",
    base_font = font_google("Orbitron")
  ),
  
  titlePanel("🌌 HoshiMap | Local Version"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("pref", "Prefecture", choices = NULL),
      sliderInput("access", "Max Travel Time (minutes)", 0, 120, 45),
      sliderInput("optical", "Max Optical Depth", 0, 5, 0.5),
      sliderInput("light", "Max Light Pollution", 0, 100, 30),
      sliderInput("sky", "Min Sky View", 0, 1, 0.5)
    ),
    
    mainPanel(
      leafletOutput("map", height = 700)
    )
  )
)

# ============================ SERVER ===========================

server <- function(input, output, session) {
  
# ========================== PREF LIST ==========================
  prefs <- gsub(
    "Accessibility_|\\.tif",
    "",
    list.files(file.path(base_path, "Accessibility_prefectures"), pattern = "\\.tif$")
  )
  
  updateSelectInput(session, "pref", choices = prefs)
  
# ======================== OBSERVATORIES ========================
  observatories <- read.csv(file.path(base_path, "Observatories.csv"), stringsAsFactors = FALSE)
  
  observatories$Longitude <- as.numeric(observatories$Longitude)
  observatories$Latitude  <- as.numeric(observatories$Latitude)
  
  obs_pal <- colorFactor(
    palette = c("#5B21B6", "#38BDF8"),
    domain = unique(observatories$Type)
  )
  
  obs_filtered <- reactive({
    req(input$pref)
    subset(observatories, Prefecture == input$pref)
  })
  
# ============================= CACHE ===========================
  raster_cache <- reactiveVal(list())
  contour_cache <- reactiveVal(list())
  
  get_data <- function(pref) {
    cache <- raster_cache()
    
    if (!pref %in% names(cache)) {
      showNotification("Loading data...", type = "message")
      
      cache[[pref]] <- list(
        landuse   = load_raster_local("Landuse_prefectures", "Landuse_", pref),
        access    = load_raster_local("Accessibility_prefectures", "Accessibility_", pref),
        opt       = load_raster_local("OpticalDepth_prefectures", "OpticalDepth_", pref),
        light     = load_raster_local("Lightpollution_prefectures", "Lightpollution_", pref),
        sky       = load_raster_local("Skyview_prefectures", "Skyview_", pref),
        hillshade = load_hillshade_local("Hillshade_prefectures", "Hillshade_", pref)
      )
      
      raster_cache(cache)
    }
    
    cache[[pref]]
  }
  
  get_contour <- function(pref) {
    cache <- contour_cache()
    
    if (!pref %in% names(cache)) {
      cache[[pref]] <- load_contour_local("Shapefile_prefectures", "Shp_", pref)
      contour_cache(cache)
    }
    
    cache[[pref]]
  }
  
# =========================== ANALYSIS ==========================
  result_raster <- reactive({
    req(input$pref)
    
    r <- get_data(input$pref)
    req(r$landuse, r$access, r$opt, r$light, r$sky)
    
    lu_mask <- landuse_mask(r$landuse)
    
    valid <- lu_mask &
      (r$access <= input$access) &
      (r$opt <= input$optical) &
      (r$light <= input$light) &
      (r$sky >= input$sky)
    
    mask(r$access, valid, maskvalues = 0)
  })
  
# ======================== VISUALIZATION ========================
  viz_rasters <- reactive({
    req(input$pref)
    
    r <- get_data(input$pref)
    res <- result_raster()
    
    hill <- r$hillshade
    if (!is.null(hill)) {
      hill <- aggregate(hill, fact = VIS_FACT, fun = mean, na.rm = TRUE)
    }
    
    res_viz <- aggregate(res, fact = VIS_FACT, fun = mean, na.rm = TRUE)
    
    list(
      result = res_viz,
      hillshade = hill
    )
  })
  
# ============================= MAP =============================
  output$map <- renderLeaflet({
    
    viz <- viz_rasters()
    contour <- get_contour(input$pref)
    obs <- obs_filtered()
    
    r <- viz$result
    hill <- viz$hillshade
    
    vals <- values(r)
    vals <- vals[!is.na(vals)]
    
    pal <- colorNumeric(
      "YlGnBu",
      domain = vals,
      na.color = "transparent"
    )
    
    m <- leaflet() %>%
      addProviderTiles("CartoDB.DarkMatter")
    
    # Hillshade
    if (!is.null(hill)) {
      m <- m %>%
        addRasterImage(hill, opacity = 0.35, project = TRUE)
    }
    
    # Raster
    m <- m %>%
      addRasterImage(r, colors = pal, opacity = 0.8, project = TRUE)
    
    # Contours
    if (!is.null(contour)) {
      m <- m %>%
        addPolygons(
          data = contour,
          color = "white",
          weight = 1.5,
          fill = FALSE
        )
    }
    
# ========================= OBSERVATORIES =======================
    if (nrow(obs) > 0) {
      m <- m %>%
        addCircleMarkers(
          data = obs,
          lng = ~Longitude,
          lat = ~Latitude,
          color = ~obs_pal(Type),
          fillColor = ~obs_pal(Type),
          fillOpacity = 0.9,
          radius = 6,
          stroke = TRUE,
          weight = 1,
          popup = ~paste0("<b>", Observatory, "</b><br>", Type)
        )
    }
    
# =========================== LEGENDS ===========================
    m <- m %>%
      addLegend(
        pal = pal,
        values = vals,
        title = "Travel Time (minutes)",
        position = "bottomleft"
      ) %>%
      addLegend(
        position = "bottomright",
        colors = c("#5B21B6", "#38BDF8"),
        labels = c("Optical / Infrared", "Radio / Astronomical"),
        title = "Observatory type"
      )
    
    m
  })
}

shinyApp(ui, server)
```
